### 2015-05-05 papers/histogram/thesis/thesis.tex

The down-going walker current $j\p{E}$ is defined in terms of the
diffusivity $\alpha\p{E}$ by
\begin{align}
  j=\alpha\f{dn_+}{dE}.
  \label{eq:walker_current_definition}
\end{align}
Computing
\begin{align}
  n\f{df}{dE}=n\sp{\f1{n}\f{dn_+}{dE}-\f{n_+}{n^2}\f{dn}{dE}}
  =\f{dn_+}{dE}-f\f{dn}{dE},
  \label{eq:walker_intermediate}
\end{align}
we can say that
\begin{align}
  \f{dn_+}{dE}=n\f{df}{dE}+f\f{dn}{dE}.
  \label{eq:walker_gradient}
\end{align}
Substituting (\ref{eq:walker_gradient}) into
(\ref{eq:walker_current_definition}) yields
\begin{align}
  j=\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
  \label{eq:walker_current_intermediate}
\end{align}
We can observe that
\begin{align}
  \f{f~dn/dE}{n~df/dE} =\p{\f{dn_+/dE-fdn/dE}{f~dn/dE}}^{-1}
  =\p{\f1f\f{dn_+}{dn}-1}^{-1}.
  \label{eq:walker_current_approx_ratio_setup}
\end{align}
Furthermore,
\begin{align}
  \p{\f1f\f{dn_+}{dn}-1}^{-1}
  =\p{\f1f\f{d\p{fn}}{dn}-1}^{-1}
  =\p{\f nf\f{df}{dn}+\f{dn}{dn}-1}^{-1}
  =\p{\f nf\f{df}{dn}}^{-1}=\f nf\f{dn}{df}=\f{d\ln n}{d\ln f},
  \label{eq:walker_current_approx_intermediate}
\end{align}
which means
\begin{align}
  \f{f~dn/dE}{n~df/dE}=\f{d\ln n}{d\ln f}.
  \label{eq:walker_current_approx_ratio}
\end{align}
The walker current is therefore
\begin{align}
  j=\alpha n\f{df}{dE}\p{1+\f{d\ln n}{d\ln f}}.
  \label{eq:walker_current_full}
\end{align}
We know that $f$ runs from $f\p{E_-}=0$ to $f\p{E_+}=1$ meaning $\ln
f$ runs from $\ln f\p{E_-}\to-\infty$ to $\ln f\p{E_+}=0$. If the
walker density $n$ varies weakly with energy, e.g. because we are
simulating with flat histogram weights, then $\ln n$ varies even
weaker still. We may therefore reasonably to assume that $\abs{d\ln
  n/d\ln f}\ll 1$ and express
\begin{align}
  j=\alpha n\f{df}{dE}.
  \label{eq:walker_current}
\end{align}
We can rearrange (\ref{eq:walker_current}) and integrate over the
energy range to say
\begin{align}
  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
  \label{eq:walker_current_integral_setup}
\end{align}
At equilibrium, the walker current $j\p{E}$ must be constant with
respect to energy. If $j\p{E}$ is not constant with respect to energy,
some energies would have either a net incurrent or net outflow of
walkers, which would mean that the ensemble is not at equilibrium. We
can therefore evaluate one of the integrals in
(\ref{eq:walker_current_integral_setup}) to get
\begin{align}
  \f1j=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
  \label{eq:walker_current_integral}
\end{align}

The optimized ensemble method minimizes the integral in
(\ref{eq:walker_current_integral}), thereby maximizing the down-going
walker current $j$, by varying the walker density $n$ with the
constraint that $n$ must remain normalized. With the additional
assumption that $\alpha\p{E}$ does not strongly depend on the weights
$w\p{E}$, the authors of \cite{optimized_ensemble} find that the
optimal walker density $n$ is
\begin{align}
  n_{\t{opt}}=\f1{\sqrt{\lambda\alpha}},
  \label{eq:optimal_walker_density}
\end{align}
where $\lambda$ is the Lagrange multiplier used in the minimization
process, and here acts merely as a normalization factor.

@@

\begin{align}
  \bk{\Delta E}=\f{\sum_{\Delta E}\Delta E~\tilde T\p{E,\Delta E}
    w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
    w\p{E+\Delta E'}/w\p{E}}
\end{align}

\begin{align}
  \bk{\Delta E^2}=\f{\sum_{\Delta E}\Delta E^2~\tilde T\p{E,\Delta E}
    w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
    w\p{E+\Delta E'}/w\p{E}}
\end{align}

\begin{align}
  \alpha=\abs{\bk{\Delta E^2}-\bk{\Delta E}^2}
\end{align}

>>>>

At equilibrium with a flat histogram, meaning $n$ is constant, the
down-going walker current $j\p{E}$ is defined in terms of the
diffusivity $\alpha\p{E}$ by
\begin{align}
  j=-\alpha\f{dn_+}{dE}.
  \label{eq:walker_current_definition}
\end{align}
This identity can be taken as the definition of the walker diffusivity
$\alpha$, which is the independent of the label we have assigned any
given walker. Substituting $n_+=fn$,
\begin{align}
  j=-\alpha\p{n\f{df}{dE}+f\f{dn}{dE}}=-\alpha n\f{df}{dE},
  \label{eq:walker_current}
\end{align}
where a constant $n$ means $dn/dE=0$. We can rearrange
(\ref{eq:walker_current}) and integrate over the energy range as
\begin{align}
  \int_{f=f\p{E_-}}^{f\p{E_+}}\f{df}{j}=-\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
  \label{eq:walker_current_integral_setup}
\end{align}
At equilibrium, the number of down-going walkers $n_+\p{E}$ at any
given energy $E$ is constant with time, which means that the walker
flux $j\p{E}$ must be constant with energy. We can therefore evaluate
one of the integrals in (\ref{eq:walker_current_integral_setup}) to
get
\begin{align}
  \f1{\abs j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
  \label{eq:walker_current_integral}
\end{align}

The optimized ensemble method minimizes the integral in
(\ref{eq:walker_current_integral}), thereby maximizing the down-going
walker current $\abs{j}$, by varying the walker density $n$ with the
constraint that $n$ must remain normalized. With the additional
assumption that $\alpha\p{E}$ does not strongly depend on the weights
$w\p{E}$, the authors of \cite{optimized_ensemble} find that the
optimal walker density $n$ is
\begin{align}
  n_{\t{opt}}\propto\f1{\sqrt{\alpha}},
  \label{eq:optimal_walker_density}
\end{align}
where the proportionality is determined by enforcing that $n$ is
normalized.

@@

By the central limit theorem, the probability distribution $P\p{E\to
  E+\Delta E}$ of Monte Carlo moves which changes the energy of a
system by $\Delta E$ is
\begin{align}
  P\p{E\to E+\Delta E}=\f1{\sqrt{2\pi\sigma^2}}
  \exp\p{-\f{\Delta E^2}{2\sigma^2}}
\end{align}
for some variance
\begin{align}
  \sigma^2=\sigma_E\p{\Delta E}^2=\bk{\Delta E^2}_E-\bk{\Delta E}_E
\end{align}
which generally depends on the energy $E$.

\vspace{3cm}

\begin{align}
  \bk{\Delta E}_E=\f{\sum_{\Delta E}\Delta E~\tilde T\p{E,\Delta E}
    w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
    w\p{E+\Delta E'}/w\p{E}}
\end{align}

\begin{align}
  \bk{\Delta E^2}_E=\f{\sum_{\Delta E}\Delta E^2~\tilde T\p{E,\Delta E}
    w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
    w\p{E+\Delta E'}/w\p{E}}
\end{align}

\begin{align}
  \alpha=\bk{\Delta E^2}_E-\bk{\Delta E}_E^2=\sigma\p{E}^2
\end{align}
